SPAGHETTI Optimizes the Quantitative Analysis of \textit{In Vitro} Phase Contrast Miscroscopy through a Generative Deep Learning Model

### 2024-11-22 rebuttal/rebuttal.tex

\pagenumbering{arabic}

\def\papertitle{Title of paper}
% \def\authors{}
% \def\journal{Some Journal}
% \def\doi{12345}
\input{rebuttal/custom_templates/AR_to_RC.tex}

\begin{refsegment}

  \AR{We thank all of the reviewers for their careful review and detailed
    suggestions to improve the manuscript. We believe that we have successfully
    addressed all comments raised by the referees from our initial submission.
    The reviewers' suggestions have significantly strengthened our manuscript
    without changing the overall story.

    To address these comments, we provided X new figure panels and included
    several new data including the following highlights:

  \begin{enumerate}
  \item Newstuff.
  \item More new stuff.
  \end{enumerate}

 Major changes to the manuscript are highlighted in orange.

}

\section{Reviewer \#1}

\RC \rlabel{rtxt:commentHere} I didn't like this.

\AR{We appreciate the reviewer noting the potential of our work and we fixed it
  like you said (\Cref{rfig:wow}). Here, we need to thank the reviewer for the
  comment (although not overdo it), then say what we need to address it. We then
  show the exact location in the text where we addressed it, along with
  additional figures. To do this, we use custom \texttt{\textbackslash{}qc},
  \texttt{\textbackslash{}qp}, and \texttt{\textbackslash{}qps} commands to copy
  (\texttt{\textbackslash{}qc}) and paste (\texttt{\textbackslash{}qp} from
  text, \texttt{\textbackslash{}qps} from supplemental). We also highlight major
  changes with \texttt{\textbackslash{}highLight} (does not work across
  paragraphs, just do another \texttt{\textbackslash{}highLight}) Check
  \texttt{paper.tex} for examples.

\qp{fixed}

\qps{otherFixInSupplemental}

\begin{rfigure}[H]
  \centering
  % Makebox here to actually center.
  \makebox[\textwidth]{\includegraphics{rebuttal/figures/figure_for_rebuttal.pdf}
  }
  {\phantomsubcaption\label{rfig:wow}}
  \caption{\panel{sub@rfig:wow}, Excerpt from \Cref{fig:astoundingOne}. Minimum
    explanation to understand figure.}
\end{rfigure}

}

\section{Reviewer \#2}

\RC The manuscript reads very well, and the quality of the language is good.

\AR{Thanks. This was noted in R\ref{rtxt:commentHere} too.}

\end{refsegment}

\printbibliography[segment=2]

\end{document}

### 2024-11-25 paper.tex
update methods section

\section*{Methods}

>>>>

\section*{Methods}
\subsection*{SPAGHETTI Model Architecture}
SPAGHETTI is built upon the orginial cyleGAN architecture for image translation \autocite{cyclegan2017}. Given a set of images from PCM ($P$) and H\&E ($H$) \todo{ensure PCM and HE are defined previously}, the goal of the model is to learn two mappings ($G_{PH}: P \rightarrow H$ and $G_{HP}: H \rightarrow P$) between a set of training images of PCM (denoted as $\{p_{i}\}^{N}_{i=1}$, where $p_{i} \epsilon P$) and H\&E (denoted as $\{h_{j}\}^{M}_{j=1}$, where $h_{j} \epsilon H$). Note that this does not require an equal number of images from both domains (ie: $N$ does not need to equal to $M$), nor require the images to be paried (ie: $p_{i}$ and $h_{j}$ are the same field of view of the same cells). To accomplished this, we empoly two adversarial discriminators, ($D_{P}$ and $D_{H}$), to distinguish between real and fake images generated by $G_{PH}$ and $G_{HP}$.

SPAGHETTI utilizes the three different loss functions implemented in the original cycleGAN paper, which are the adversarial loss ($\mathcal{L}_{GAN}$), the cycle-consistency loss ($\mathcal{L}_{cyc}$), and the identity loss ($\mathcal{L}_{idt}$). The adversarial loss is used to train the generators to generate realistic images, while the cycle-consistency loss is used to ensure that the images generated after two generators are consistent with the input images. The identity loss is used to ensure that the generators do not alter the input images if the images are passed through the generator from the same domain. Those three loss functions are defined as follows:

\begin{equation}
  \begin{aligned}
  \mathcal{L}_{GAN}(G_{PH}, D_{H}, P, H) = \mathbb{E}_{h \epsilon H}[\log D_{H}(h)] + \mathbb{E}_{p \epsilon P}[\log(1 - D_{H}(G_{PH}(p)))] \\
  \mathcal{L}_{GAN}(G_{HP}, D_{P}, H, P) = \mathbb{E}_{p \epsilon P}[\log D_{P}(p)] + \mathbb{E}_{h \epsilon H}[\log(1 - D_{P}(G_{HP}(h)))] \\
  \mathcal{L}_{cyc}(G_{PH}, G_{HP}) = \mathbb{E}_{p \epsilon P}[\|G_{HP}(G_{PH}(p)) - p\|_{1}] + \mathbb{E}_{h \epsilon H}[\|G_{PH}(G_{HP}(h)) - h\|_{1}] \\
  \mathcal{L}_{idt}(G_{PH}, G_{HP}) = \mathbb{E}_{p \epsilon P}[\|G_{HP}(p) - p\|_{1}] + \mathbb{E}_{h \epsilon H}[\|G_{PH}(h) - h\|_{1}] \\
  \end{aligned}
\end{equation}

In addition, to solve the issues of hallusination, which is a major concern in medical image translation \autocite{hallucination}, we introduced a new loss function called the SSIM loss ($\mathcal{L}_{SSIM}$) \todo{ensure SSIM is defined properly before}. The SSIM loss is used to penalize the generator if the generated images are not consistent with the input images by introducing or neglecting additional features. The SSIM loss is calculated from the SSIM value, which uses a sliding window approach and the SSIM value is calculated for each window and averaged across the entire image. The SSIM value is a measure of the similarity between two images, which ranges from -1 to 1, where 1 indicates that the two images are identical \autocite{ssim}. It is calculated based on the luminance, contrast, and structure of the images $x$ and $y$ as follows:

\begin{equation}
  \begin{aligned}
  SSIM(x, y) = \frac{(2\mu_{x}\mu_{y} + c_{1})(2\sigma_{xy} + c_{2})}{(\mu_{x}^{2} + \mu_{y}^{2} + c_{1})(\sigma_{x}^{2} + \sigma_{y}^{2} + c_{2})}
  \end{aligned}
\end{equation}

where $\mu_{x}$ and $\mu_{y}$ are the average of $x$ and $y$, $\sigma_{x}^{2}$ and $\sigma_{y}^{2}$ are the variance of $x$ and $y$, $\sigma_{xy}$ is the covariance of $x$ and $y$, and $c_{1}$ and $c_{2}$ are constants to stabilize the division with weak denominator. The SSIM loss is then defined as follows:

\begin{equation}
  \begin{aligned}
  \mathcal{L}_{SSIM}(G_{PH}, G_{HP}) = \frac{1}{N} \sum_{i=1}^{N} \frac{(1 - SSIM(G_{PH}(p_{i}), p_{i}))}{2} + \frac{1}{M} \sum_{j=1}^{M} \frac{(1 - SSIM(G_{HP}(h_{j}), h_{j}))}{2}
  \end{aligned}
\end{equation}

The overall loss is a combination of all the above four losses, which is defined as follows:

\begin{equation}
  \begin{aligned}
    \mathcal{L}(G_{PH}, G_{HP}, D_{P}, D_{H}) = & \mathcal{L}_{GAN}(G_{PH}, D_{H}, P, H) \\
    & + \mathcal{L}_{GAN}(G_{HP}, D_{P}, H, P) \\
    & + \lambda_{cyc}\mathcal{L}_{cyc}(G_{PH}, G_{HP}) \\
    & + \lambda_{idt}\mathcal{L}_{idt}(G_{PH}, G_{HP}) \\
    & + \lambda_{SSIM}\mathcal{L}_{SSIM}(G_{PH}, G_{HP})
  \end{aligned}
\end{equation}

where $\lambda_{cyc}$, $\lambda_{idt}$, and $\lambda_{SSIM}$ are the hyperparameters that control the relative importance of the cycle-consistency, identity, and SSIM loss terms respectively. The overall objective function is optimized by solving the following min-max problem, where the generators are optimized to minimize the loss function, while the discriminators are optimized to maximize the loss function:

\begin{equation}
  \begin{aligned}
    \min_{G_{PH}, G_{HP}} \max_{D_{P}, D_{H}} \mathcal{L}(G_{PH}, G_{HP}, D_{P}, D_{H})
  \end{aligned}
\end{equation}

\subsection*{Training Details}
We trained SPAGHETTI and an original cycleGAN architecture using the LIVECell dataset \autocite{LIVECell} for PCM (N = 5326) and the PanNuke dataset \autocite{pannuke} for H\&E (M = 7904). The RGB images of PCM were cropped randomly to a quarter of its original size while the H\&E images were kept as is, and all images were resized to 256x256 pixels and converted to the range of [0, 1]. They were then normalized to a mean of 0.5 and a standard deviation of 0.5. The generators were implemented using a ResNet architecture \autocite{resnet} with 9 residual blocks. We split the dataset to 80\% training and 20\% testing, before using a batch size of 16 and the AdamW optimizer \autocite{adamw} with a learning rate of 0.001 and a weight decay of 0.01 for training. The hyperparameters $\lambda_{cyc}$ and $\lambda_{idt}$ were set to 10.0 and 5.0 for both SPAGHETTI and the original cycleGAN architecture, and $\lambda_{SSIM}$ was set to 3.0 for SPAGHETTI. Both models were trained for 100 epochs on 8 NVIDIA T4 Tensor Core GPUs on Vector Institute's Vaughan HPC. After the training was completed, we utilized the PCM to H\&E generator ($G_{PH}$) as the generators for SPAGHETTI and the original cycleGAN to generate H\&E-like images from PCM images.

\subsection*{LIVECell Feature Attention Visualization and Quantification}
We visualized the attention map of the \textit{CLS} token, which is utilized as the feature of image \autocite{vit}, from the last transformer layer of Phikon-v2, an open source foundational feature extractor designed for H\&E images \autocite{phikonv2}, on the 20\% testing portion of the LIVECell dataset. We reshaped the \textit{CLS} token to be the shape of 16x16 pixels, before extrapolating it back bicubicly to the original image size of 256x256 pixels. We then computed the attention map by averaging the attention weights across all heads and normalizing the attention weights to the range of [0, 1]. We finally overlaid the attention map on the original image to visualize the feature attention by utilizing a plasma colour map.

To quantify whether the transformation allows the model to focus better on the cell populations of the images, we took the top 60\% of the attention score as the predicted cell masks and calculated the Dice score with the ground truth cell masks. The Dice score is defined as follows by comparing the ground truth cell masks ($A$) and the predicted cell masks ($B$):

\begin{equation} \label{eq:dice}
  \begin{aligned}
    Dice = \frac{2|A \cap B|}{|A| + |B|}
  \end{aligned}
\end{equation}

We repeated the above process for using SPAGHETTI as the pre-processing step, using the original cycleGAN architecture as the pre-processing step, and using the original PCM images before feeding the images into Phikon-v2. We performed two-tailed paired t-tests to compare the Dice scores between the different pre-processing steps.

\subsection*{LIVECell Cell Segmentation}
We performed cell segmnetaions using the two popular existing cell segmentors CellPose \autocite{cellpose} and StarDist \autocite{stardist} on the 20\% testing portion of the LIVECell dataset. In particular, we used the tissuenet\_cp3 model for CellPose and the 2D\_versatile\_he model for StarDist. We reported the segmentation results using the Dice coefficient defined in equation (\ref{eq:dice}). We repeated the above process for using SPAGHETTI as the pre-processing step, using the original cycleGAN architecture as the pre-processing step, and using the original PCM images before feeding the images into the cell segmentors. As a control for CellPose, we also pass the original PCM images through the cyto3 model of CellPose. We performed two-tailed paired t-tests to compare the Dice scores between the different pre-processing steps.

\subsection*{LIVECell Cell Type Classification}

\subsection*{C2C12 Cell Media Condition Classification}
